#  IM Metadata Specification {#meta}

```{r set up, include=FALSE}
load("./hub.RData")
#source("/Users/isaiahnyabuto/Documents/Workspace/utils.R")
library(knitr)
library(kableExtra)
library(dplyr)
#library(plyr)


# baseurl <- "https://im-dev.psi-mis.org/"
# base <- substr(baseurl,9,26)
# usr <- keyringr::get_kc_account(base,type = "internet")
# pwd <- keyringr::decrypt_kc_pw(base, type = "internet")
# 
# loginDHIS2(baseurl,usr,pwd)
# 
# ou_group_set <- function(){
#   r <- GET(paste0(baseurl,"api/organisationUnitGroupSets?fields=name,organisationUnitGroups[name]"))
# 
#   set <- fromJSON(content(r,"text"))$organisationUnitGroupSets
# 
#   if (!is.null(set)){
#     # covert df to a string
#     set$organisationUnitGroups <- substring(as.character(set$organisationUnitGroups), regexpr("['\"']",
#                                                                                     as.character(set$organisationUnitGroups))+1)
# 
#     set$organisationUnitGroups <- str_remove_all(set$organisationUnitGroups,"[()\"']")
# 
#     names(set) <- c("OU group set", "OU groups")
#     set
#   }else{
#     ""
#   }
# 
# 
# }
# 
# categories <- function(){
#   r <- GET(paste0(baseurl, "api/categories?fields=name,id,dataDimensionType,categoryOptions[name]&filter=name:like:CORE"))
#   
#   cat <- fromJSON(content(r,"text"))$categories
#   
#   if (!is.null(cat)){
#     # df to string
#     
#     cat$categoryOptions <- substring(as.character(cat$categoryOptions), regexpr("['\"']",
#                                                                                 as.character(cat$categoryOptions))+1)
#     cat$categoryOptions <- str_remove_all(cat$categoryOptions, "[()\"']")
#     
#     cat
#     
#   }else{
#     ""
#   }
# }
# 
# category_combo <- function(){
#   r <- GET(paste0(baseurl, "api/categoryCombos?fields=name,id,dataDimensionType,categories[name]&filter=name:like:CORE"))
#   
#   cat <- fromJSON(content(r,"text"))$categoryCombos
#   
#   if (!is.null(cat)){
#     # df to string
#     
#     cat$categories <- substring(as.character(cat$categories), regexpr("['\"']",
#                                                                                 as.character(cat$categories))+1)
#     cat$categories <- str_remove_all(cat$categories, "[()\"']")
#     
#     cat
#     
#   }else{
#     ""
#   }
# }

```

## Introduction

In the previous chapter, we learned about the IM PMP indicators and how they are stored or constructed in the IM Data Hub. That is essential knowledge for you to start monitoring IM PMP indicators from the source to how they are presented in the IM Data Hub.

In this chapter, we'll dive deep to the bottom and explore what lies beneath the IM Data Hub, the metadata. Metadata defines what, why, where, and how parts are set up in the IM Data Hub. They provide references to different objects in the IM Data Hub.

Already you are familiar with data sets, data elements, indicators: all these are essentially metadata themselves. 

The IM Data Hub metadata are specified in either CSV, JSON or XML formats and are available at the metadata API endpoint:

`https://imdatahub.org/api/metadata`

Metadata has a name, uid, code that uniquely identifies the object in the IM Data Hub. It also has a sharing setting that defines the access group (have a look to the section 'Access model' for further info on this). 

All IM Data Hub metadata can be accessed individually by specifying their names or ids on the API endpoint. For example, the metadata for a data element with the short name `suspected malaria cases` is accessed at:

`https://imdatahub.org/api/29/dataElements/eHWYhy8vO8q.json`

In the following sections we discuss the IM Data Hub metadata in three main groups:

1. Organization Unit stores information about the location where IM indicator data is reported, analyzed, or presented, 
2. Metadata that specifies what, why, and how IM Data Hub components are set up. They include; datasets, data elements, sections, data elements groups and group sets, organization unit groups and group sets, indicators, indicator groups, categories, category options, categories, category combinations, and category option combinations,
3. CORE metadata - any metadata that is shared publicly within the IM Data Hub for re-usability across the IM Data Hub.


## Organization Units (OUs)

Organization Units specify the location or where the IM data elements are collected, analyzed, or presented. 

There are two main types of OUs in the IM Data Hub:

1. Test - a list of OUs for development and testing objects. This is where we have been doing our practices,
2. World - a list of OUs for production or real use in the IM Data Hub.

The OUs in DHIS2 are structured in the form of a tree or through a hierarchy to allow entries, data processing, and drill downs/ups on the Analytics apps. They also define access levels in the IM Data Hub (have a look to the section 'Access model' for further info on this)).

The IM Data Hub OUs are grouped into OU groups and group sets for disaggregation of IM data elements and indicators. We discuss the groups in section \@ref{ougroup}

### OU Trees

As we mentioned above, the IM Data Hub OUs are organized in a tree-like structure that provides different levels for reporting, data processing, and presentation of IM data. In this section, we describe the OU structure by country.

The structure is in the form hierarchy and begins with the country to subnational level 4.

Country| Subnational level 1| Subnational level 2| Subnational level 3| Subnational level 4 |
-------|:-------------------|:-------------------|:-------------------|:--------------------|
Cameroon| Region |District| Aire de santé| Facility
Cote d’Ivoire| Region| District| Facility| /
DRC| Province| Zone de santé | Aire de santé | Facility
Kenya | County | Sub-county | Ward | Facility
Ghana | Region | District | Sub-district | Facility 
Mali | Region | District | Aire de santé |Facility
Niger | Province | District Sanitaire | Aire de santé | Facility 
Sierra Leone | Province | District | Chiefdom | Facility
Zambia | Province | District | Facility | /

The following is the Metadata endpoint for organization unit levels.
`https://imdatahub.org/api/organisationUnitLevels.json`

Next we’ll talk about the OU groups and group sets. 

### OU Groups, OU Group sets

The IM Data Hub OUs consist of the following OU groups and group sets by countries and the type or owenrship of facility: 

```{r, echo=FALSE}
load("./hub.RData")
#ou_groups <- data.frame(`Country (# of units)` = c("CD (<b>114</b>):", "CI (<b>777</b>):", "CM (<b>751</b>):",
#                                                   "KE  (<b>1775</b>):", "GH (<b>7,948</b>):", "ML (<b>410</b>):",
#"SL (<b>303</b>):", "ZM (<b>227</b>)"))

# ou_group <- ou_group_set()
# save(ou_group, file = "./hub.RData")

# resave(ou_group,file = "./hub.RData")
kable(ou_group, caption = "A table showing OU groups and group set") %>% kable_styling()
```


## Metadata

The IM Data Hub implements the following metadata types:

```{r, echo=FALSE}

metadata_types <- data.frame(`Metadata` = c("Data elements","Date element groups","Data sets",
                                                 "Sections","Category combos","Periods", "Indicator", "Indicator groups"),
                             `Description` = c("<h3>Description</h3></p>Data elements forms the basis of the IM Data Hub. They define what is recorded in the system.</p></p>IM Data Hub data elements follows PSI DHIS2 standards for names and short names & the IM Indicator matrix for the form names.</p><p>For instance: <em>IM CR - DX Suspected malaria cases</em> is a data element that records suspected malaria cases</p>
",
"<h3>Description</h3><p>Data element groups provides a mechanism for classifying IM Data Hub data elements into a common theme.</p></p>For instance: <em>CD Supportive Supervision</em> groups together all IM supportive supervision data elements. recorded in Congo</p>",
"<h3>Description</h3></p>Data reporting in the IM Data Hub is organized through the use of the use of data sets.</p><p> A data set is a collection of data elements grouped together for data collection.</p><p> For instance: <em>IM Case Reporting</em> is a data set used to report case reporting data globally</p>
",
"<h3>Description</h3><p>Section are used to split the IM Data Hub data sets into small reportable chunks known as modules.</p> <p>They allow for a bit more flexibility when it comes to using the tabular forms in DHIS2 and are quick and simple to design.</p><p> For instance: <em>Diagnosis, Treatment, MIP and Reporting</em> are section within the IM Case Reporting data set</p>",
"<h3>Description</h3><p>Category combos are used to apply IM disaggregation onto the data elements and indicators.They allow multiple categories to be combined into a related set.</p><p> For instance: <em>IM PMP DX 1  - Percentage of confirmed malaria cases (NP)</em> is disaggregated according to the following categories; <em>CORE - Sex | CORE - Parasitological test | CORE - Age-group</em>
",
"<h3>Description</h3><p>IM Data Hub supports DHIS2 standard period types to report and analyze performance over time.</p><p> For instance: <em>201901</em> to report on January of 2019 </p>
",
"<h3>Description</h3><p>Indicators allows the computation of IM data. They consist of a numerator, a denominator and a data element</p><p>",

"<h3>Description</h3><p>Similar to data elements groups, indicator groups - groups together IM indicator to allow easy access from the analytical tools </p><p> For instance: <em>KE PMP</em> is an indicator group with all PMP indicators for Kenya </p>"),

stringsAsFactors = F)

kable(metadata_types, caption = "A table showing metadata types implimented in the IM Data Hub", escape = F) %>% kable_styling()
```


## CORE Metadata

CORE Metadata are reserved objects in the data hub. They are shared publicly within the IM Data Hub for re-use in the IM Data Hub. There are two main types of `CORE metadata.`

1. Category: a group of cat options for disaggregation/attributes,
2. Category Combination: a combination of classes used to apply disaggregation onto data elements/indicators. 

The following is a list of all IM Data Hub CORE metadata.

### Category

```{r, echo=FALSE}

# cat <- categories()
# resave(cat, file = "./hub.RData")

kable(cat, caption = "A table showing core categories set in the IM Data Hub", escape = F) %>% kable_styling()
```

### Category Combination 

```{r, echo=FALSE}

# cat_combos <- category_combo()
# 
# resave(cat_combos, file = "./hub.RData")

kable(cat_combos, caption = "A table showing core category combos set in the IM Data Hub", escape = F) %>% kable_styling()
```
